1.3 Meet R - R Practice

### 2019-09-03 content/class/03-class.Rmd
Update class 3 materials

- [<i class="fas fa-external-link-square-alt"></i> Lecture Slides](/slides/02-slides.html)

# Appendix

@@

@@

@@

One of the best/worst things about `R` is that it is a language, meaning there are multiple ways that you can accomplish the same task.

@@

- `:` creates a series of integers

@@

- `rep(., times =)`^[`.` is a placeholder here.] repeats an element a specified number of times

@@

- We can combine these:

@@

>>>>

- [<i class="fas fa-external-link-square-alt"></i> Lecture Slides](/slides/03-slides.html)

# Practice Problems

Today you will be working on [R practice problems](/r/03-r-practice.html). Answers will be posted later.

# Statistics and Software Survey

Please turn in your math surveys. I will post "answers" after both sections have turned in their surveys.

# Additional Useful Information

@@

We will be using the trading room computers all semester, and I will ensure you have access to the computers outside of class so that you can work on homeworks and your projects on them. However, since R is free, I **strongly recommend** you download and install it on your own computers. You may decide also you prefer to use your own computers in class when we work with R.

@@

## Other Resources for R

My R4EH guide.

@@

One of the best/worst things about `R` is that it is a language, meaning there are multiple ways that you can accomplish the same task. Here are a few alternative methods relevant to what we have learned so far that might prove useful.

@@

1. `:` creates a series of integers

@@

2. `rep(., times =)`^[The `.` is a placeholder here.] repeats an element a specified number of times

@@

3. We can combine these:

@@

### Other Object Types

### 2019-09-03 content/r/03-r-practice.Rmd
Update class 3 materials

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(pander)
source(file.path(here::here(), "lib", "graphics.R"))
```

>>>>

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(pander)
source(file.path(here::here(), "lib", "graphics.R"))
```
- [<i class="fas fa-file-pdf"></i> `Download as PDF`](/r/03-r-practice.pdf)

### 2019-09-03 static/r/03-r-practice.Rmd
Update class 3 materials

## Getting Set Up

Before we begin, start a new file with \texttt{File} $\rightarrow$ \texttt{New File} $\rightarrow$ \texttt{R Script}. As you work through this sheet in the console in `R`, also add (copy/paste) your commands that work into this new file. At the end, save it, and run to execute all of your commands at once.

## Creating Objects

### 1. Create a vector called "me" with two objects, your first name, and your last name. Then call the vector to inspect it. Confirm it is a character class vector.

### 2. Create a vector called "x" with all the even integers from 2 to 10.

### 3. Find the mean of x with `mean()`

### 4. Now take the following pdf of random variable $Y$:

\begin{table}[h!]
\centering
  \begin{tabular}{ll}
  $y_i$ & $p_i$\\ \hline
  2 & 0.50\\
  4 & 0.25\\
  6 & 0.25\\ \hline
  \end{tabular}
\end{table}

Calculate the standard deviation "manually" using our table method. You can look at the source code of Lecture 4 for my example.

### a. Creating two vectors, one called $y.i$ and one called $p.i$, with the data above.

### b. Merge them into a data frame called $rv$ with `data.frame(y.i,p.i)`. Call `rv` to inspect it.

### c. Find the expected value of $Y$ by taking the sum of each value of `y.i` multiplied by `p.i` with the `sum()` command.

### d. Creating a new column in `rv` called `deviations`, where you subtract the mean from each `y.i` value. Call `rv` again to make sure it's now there.

### e. Create another column in `rv` called `devsq`, where you square the deviations from part d. Call `rv` again to make sure it's now there.

### f. Now add another column in `rv` called `weighteddevsq`, where you multiply the squared deviations in part e. by the associated probability `p.i`. Call `rv` again to make sure it's now there.

### g. Finally, take the sum of `weighteddevsq` to get variance. Square root this to get standard deviation.

### 5. The mean height of adults is 65 inches, with a standard deviation of 4 inches. Use the normal distribution to find the probabilities of the following scenarios:

### a. Find the probability of someone being *at least* 60 inches tall using `pnorm()`.

### b. Find the probability of someone being *at most* 60 inches tall.

### c. Find the probability of someone being between 61 and 69 inches tall. Why is this number familiar?

### d. Find the probability of someone being between 57 and 73 inches tall. Why is this number familiar?

## Playing with a Data Set

For the following questions, use the \texttt{diamonds} dataset, included as part of `ggplot2`.

### 6. Install `ggplot2`

### 7. Load `ggplot2` with the `library()` command

### 8. Get the structure of the diamonds data.frame. What are the different variables and what kind of data does each contain?

### 9. Get summary statistics for `carat`, `depth`, `table`, and `price`

@@

### 11. Now rerun the `summary()` command on the entire data frame

### 12. Plot a histogram of price.

### 13. Plot a boxplot of price by diamond color.

## Execute your R Script

>>>>

# Getting Set Up

Before we begin, start a new file with `File` $\rightarrow$ `New File` $\rightarrow$ `R Script`. As you work through this sheet in the console in `R`, also add (copy/paste) your commands that work into this new file. At the end, save it, and run to execute all of your commands at once.

# Creating Objects

## 1.

a. Create a vector called `me` with two objects, your first name, and your last name.
b. Call the vector to inspect it.
c. Confirm it is a character class vector.

## 2. Use R's help functions to determine what the `paste()` function does. Then paste together your name from `me`.

## 3. Create a vector called `my_vector` with all the even integers from 2 to 10.

## 4. Find the mean of x with `mean()`.

## 5. Add up all the numbers from 18 to 763, then get the mean

\clearpage

# Playing with Data

For the following questions, we will the `diamonds` dataset, included as part of `ggplot2`.

### 6. Install `ggplot2`.

### 7. Load `ggplot2` with the `library()` command.

### 8. Get the `str`ucture of the diamonds data frame. What are the different variables and what kind of data does each contain?

### 9. Get summary statistics separately for `carat`, `depth`, `table`, and `price`.

@@

### 11. Now rerun the `summary()` command on the entire data frame.

### 12. Now look only at (subset) the first 4 diamonds in the dataset.

### 13. Now look only at (subset) the third and seventh diamond in the dataset.

### 14. Now look only at (subset) the second column of the dataset.

### 15. Do this again, but look using the `$` to pull up the second column by name.

### 16. Now look only at diamonds that have a `carat` greater than or equal to 1.

### 17. Now look only at diamonds that have a VVS1 `clarity`.

### 17. Now look only at dimaonds that have a color of E, F, I, and J.

### 18. Now look only at diamonds that have a `carat` greater than or equal to 1 and a VVS1 `clarity`.

### 19. Get the average price of diamonds in question 18.^[Hints: use your subset command as an argument to the mean function. You will not need a comma here because you are looking for a single row.]

### 20. What is the highest price for a diamond with a 1.0 carat, D color, and VVS1 clarity?

# Execute your R Script
